Users Guide for Corrfunc

### 2015-01-28 paper/users-guide.tex
Changed variables to be consistent across codes

\usepackage{geometry}
\usepackage{amsmath}
\usepackage{aas_macros}
\usepackage{rotating}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{color}
\usepackage{listings}
\usepackage{caption}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue,breaklinks=true]{hyperref}
\usepackage{dirtree}
\usepackage{adjustbox}
\usepackage{bytefield}
\usepackage{chngcntr}
\usepackage{alltt}

@@

%% Some of the short-cuts I use
\newcommand{\rmax}{\ensuremath{{r_{max}}}\xspace}
\newcommand{\xir}{\ensuremath{{\xi(r)}}\xspace}
\newcommand{\wprp}{\ensuremath{{w_p(r_p)}}\xspace}
\newcommand{\xirppi}{\ensuremath{{\xi(r_p,\pi)}}\xspace}
\newcommand{\pimax}{\ensuremath{{\pi_{\rm max}}}\xspace}
\newcommand{\hMpc}{\ensuremath{{h^{-1}Mpc}\xspace}}
\newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}
\newcommand{\clang}{{\texttt{clang}}\xspace}
\newcommand{\icc}{{\texttt{icc}}\xspace}
\newcommand{\gcc}{{\texttt{gcc}}\xspace}

@@

\title{User Guide for Corrfunc}
%% Abuse the author macro to show all the info I want.
\author{Manodeep Sinha, \\
  Department of Physics \& Astronomy,\\
  Vanderbilt University,\\
  Nashville, TN 37235.\\
  \href{mailto:manodeep@gmail.com}{\texttt{manodeep@gmail.com}}
}

@@

\stdsection{Running the Codes}
The codes should run straight out of the box. Note, the \xir and \xirppi can compute both auto and cross-correlations with and without PERIODIC boundary conditions whereas the \wprp code 
only computes the auto-correlation with PERIODIC boundary conditions. Also, since both the \xir and \xirppi codes compute cross-correlations, pairs are {\em double-counted} in both 
\xir and \xirppi codes. You could, in theory, make the auto-correlation bit faster by only computing unique pairs (as is done in \wprp, see Section~\ref{section:pair_counting_wp} 
for details). However, since the total run-time will dominated by the cross-correlation (the number of randoms is typically an order of magnitude larger than the number 
of data points), I have not implemented those optimizations for the auto-correlation calculations. 
\subsection{Input Files}
The codes currently can handle these types of input data files:
\begin{itemize}
\item \texttt{ascii} -- White-space separated columns, format code is `a'.
\item \texttt{csv}   -- Comma-separated values, format code is `c'.
\item \texttt{fast-food} -- Fast-food, fortran binary format, format code is `f'.
The fast-food file format is described in detail in Section~\ref{section:fastfood}.
\end{itemize}
For the \texttt{ascii} and \texttt{csv} files, the code reads in the first three columns as the co-moving \texttt{X/Y/Z} arrays. Note, that more columns 
can be present but the code will ignore those columns.

@@

\flushleft The \texttt{fast-food} file consists of a header:
\begin{lstlisting}[caption={The header format for \texttt{fast-food} files}]
int idat[5];
float fdat[9];
float znow;
\end{lstlisting}
For the purposes of these correlation function codes, the only useful quantity is \texttt{idat[1]} which contains \texttt{N} -- the number of particles 
in the data file.

@@

\vspace{0.2in}
\begin{minipage}[t]{\textwidth}
\centerline{\underline{\bfseries Byte-structure of the X/Y/Z arrays in a fast-food file}}
\vspace{0.2in}
\begin{minipage}[t]{0.5\textwidth}
\centering
\textbf{floats}\\[3ex]
\begin{bytefield}{4}
  \bitheader{0-3} \\
  \bitbox{4}{\texttt{N*4}} \\
  \wordbox{1}{X[0]} \\
  \wordbox{1}{X[1]} \\
  \wordbox[]{1}{$\vdots$} \\[1ex]
  \wordbox{1}{X[N-1]} \\
  \bitbox{4}{\texttt{N*4}}
\end{bytefield} 
\end{minipage}%
\begin{minipage}[t]{0.5\textwidth}
\centering
\textbf{doubles}\\[3ex]
\begin{bytefield}{8}
  \bitheader{0-7} \\
  \bitbox{4}{\texttt{N*8}} \\
  \wordbox{1}{X[0]} \\
  \wordbox{1}{X[1]} \\
  \wordbox[]{1}{$\vdots$} \\[1ex]
  \wordbox{1}{X[N-1]} \\
  \bitbox{4}{\texttt{N*8}}
\end{bytefield}
\end{minipage}
\end{minipage}
\subsection{Specifying the radial bins}\label{section:bins}
The codes were intentionaly designed to read in a set of (somewhat) arbitrary\footnote{I assume bins are non-overlapping} set of bin specifications from a file. This way, you can specify disjoint bin-edges as well as 
use $0.0$ as a bin edge (which would be impossible if $\log$ bins are assumed). The bins are to be specified in a (white-space separated) text file in this manner:
\begin{alltt}
r\_low[0]  r\_high[0]
r\_low[1]  r\_high[1]
\vdots
r\_low[nbins-1]  r\_high[nbins-1]
\end{alltt}
where, \texttt{r\_low[i]} and \texttt{r\_high[i]} are the left and right edges of the \texttt{i}'th bin respectively. The text files should contains as many lines as the number of 
bins desired. The \texttt{logbins} executable can be used to create such a text file containing $\log$ bins. The syntax for running \texttt{logbins} is:\\
\vspace{0.2in}
\texttt{./logbins <rmin> <rmax> <nbins>   > <text filename>}
\vspace{0.2in}
\\An example of such a file with radial bins is the file \texttt{bins} in the \texttt{tests} directory. 
{\em Note, all of the three codes print the correlation function to \texttt{stdout} -- so be sure to redirect \texttt{stdout} to an output file. }

\subsection{Running \texorpdfstring{\xir}{xi(r)}}
To run the correlation function in 3-D, you will need to run the executable \texttt{DD} from either the \texttt{bin} or \texttt{xi\_of\_r} directory. Note, that 
\xir double-counts the pairs (as does \xirppi). The inputs to \texttt{DD} are:
\begin{itemize}
\item \texttt{file1}   -- the file name for the first file.
\item \texttt{format1} -- the file format for the first file.
\item \texttt{file2}   -- the file name for the second file.
\item \texttt{format2} -- the file format for the second file.
\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
\end{itemize}
Thus, the code can compute an auto-correlation (when \texttt{file1,format1} and \texttt{file2,format2} are identical) or a cross-correlation (when \texttt{file1} and \texttt{file2} are different). Out of the box, 
some sample \texttt{DD} call directory) can be:
\begin{itemize}
\item auto-correlation --  {\scriptsize \texttt{./DD  ../tests/data/gals\_Mr19.ff f ../tests/data/gals\_Mr19.ff f ../tests/bins 4 > Mr19\_output.DD}}
\item cross-correlation -- {\scriptsize \texttt{./DD ../tests/data/cmassmock\_Zspace.ff f ../tests/data/random\_Zspace.ff f ../tests/bins 4 > cmass\_output.DR}}
\end{itemize}
The output (printed to stdout) has \texttt{nbins} rows; each row contains the columns \texttt{<Npairs> <rpavg> <rmin> <rmax>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
Makefile option OUTPUT\_RPAVG has been enabled. Here, each line of the output represents the \texttt{i}'th radial bin.

\subsection{Running \texorpdfstring{\xirppi}{xi(rp,pi)}}
To run the correlation function in 2-D for \xirppi, you will need to run the executable \texttt{DDrppi} from either the \texttt{bin} or \texttt{xi\_rp\_pi} directory. Note, that 
\xirppi double-counts the pairs (as does \xir).The inputs 
to \texttt{DDrppi} are:
\begin{itemize}
\item \texttt{file1}   -- the file name for the first file.
\item \texttt{format1} -- the file format for the first file.
\item \texttt{file2}   -- the file name for the second file.
\item \texttt{format2} -- the file format for the second file.
\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
\item \texttt{pimax}   -- the maximum distance to consider in the $\pi$ direction. The code uses \texttt{Z} axis as the $\pi$ (line-of-sight) direction. 
\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
\end{itemize}
Thus, the code can compute an auto-correlation (when \texttt{file1,format1} and \texttt{file2,format2} are identical) or a cross-correlation (when \texttt{file1} and \texttt{file2} are different). Out of the box, 
some sample \texttt{DDrppi} call directory) can be:
\begin{itemize}
\item auto-correlation --  {\scriptsize \texttt{./DDrppi  ../tests/data/gals\_Mr19.ff f ../tests/data/gals\_Mr19.ff f ../tests/bins 40.0 4 > Mr19\_output\_rppi.DD}}
\item cross-correlation -- {\scriptsize \texttt{./DDrppi ../tests/data/cmassmock\_Zspace.ff f ../tests/data/random\_Zspace.ff f ../tests/bins 80.0 4 > cmass\_output\_rppi.DR}}
\end{itemize}
The output (printed to stdout) has \texttt{nbins} rows; each row contains the columns \texttt{<Npairs> <rpavg> <log(rmin)> <pi\_upper>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
Makefile option OUTPUT\_RPAVG has been enabled. The code bins in 1 Mpc/h in the $\pi$ direction by default. Therefore, 
the total number of bins in the output file will be \texttt{nbins$\times$\pimax}, where each radial bin is further broken into \pimax bins along the 
$\pi$ direction.

@@

\subsection{Running \texorpdfstring{\wprp}{wp(rp)}}
To run the projected correlation function, you will need to run the executable \texttt{wp} from either the \texttt{bin} or \texttt{wp} directory. The inputs 
to \texttt{wp} are:
\begin{itemize}
\item \texttt{boxsize} -- the boxsize for the periodic box. 
\item \texttt{file}    -- file name.
\item \texttt{format}  -- file format.
\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
\item \texttt{pimax}   -- the maximum distance to consider in the $\pi$ direction. The code uses \texttt{Z} axis as the $\pi$ (line-of-sight) direction. 
\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
\end{itemize}
The \texttt{wp} code only computes an auto-correlation with PERIODIC boundary conditions, irrespective of the options set in \texttt{common.mk}. 
\begin{itemize}
\item {\scriptsize \texttt{./wp  420.0 ../tests/data/gals\_Mr19.ff f ../tests/bins 40.0 4 > Mr19\_output.wp}}
\end{itemize}
The output (printed to stdout) has \texttt{nbins} rows; each row contains the columns \texttt{<wp> <rpavg> <rmin> <rmax> <Npairs>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
Makefile option OUTPUT\_RPAVG has been enabled.

@@

\subsection{Partitioning the Particles based on \texorpdfstring{\rmax}{rmax}}
\begin{figure}
\centering
\includegraphics[width=\textwidth,clip=true]{tikz_grid}
\caption{A 2-D grid showing the bin-lattice partitioning scheme. The bigger square show the entire
domain, the red circles show a random distribution of 100 particles. Let's say we want to compute all pairs
for the target blue point, then we would only have to consider red points that are within one cell (the dark shaded region).
A circle with radius \rmax is also drawn to shown the actual pairs that will eventually count in the correlation function.}
\label{fig:grid}
\end{figure}

@@

By wrapping the elements if the first data-set, we can avoid these operations in the following \texttt{j-loop} over all particles in \texttt{second}.

\begin{lstlisting}[label={code:looping_over_second_particles},caption={AVX intrinsics for looping over all particles in the \texttt{second} cell. PERIODIC boundary conditions have 
already been accounted for in \texttt{x1pos,y1pos,z1pos} variables.}]
const AVX_FLOATS m_x1pos = AVX_SET_FLOAT(x1pos);
const AVX_FLOATS m_y1pos = AVX_SET_FLOAT(y1pos);
const AVX_FLOATS m_z1pos = AVX_SET_FLOAT(z1pos);

@@

The three codes \xir, \xirppi and \wprp diverge somewhat after this point. For \xir, we need to calculate the full 3-D separation, whereas for \xirppi and \wprp 
the separation is the projected distance. I will discuss further implementations in the following sub-sections dedicated to each code. 
\subsubsection{Pair-counting for \texorpdfstring{\xir}{xi(r)}}\label{section:pair_counting_r}
Pair-counting for \xir is the most straight-forward. Once all the \texttt{NVEC} separations have been computed, we use bit-masks to check if any separations fall 
within the range \texttt{sqr\_rpmin} and \texttt{sqr\_rpmax}. If not, we \texttt{continue} with the \texttt{j-loop}. 
\begin{lstlisting}[label={code:masks_xi},caption={Bit-masks in \xir.}]
m_mask_left = AVX_COMPARE_FLOATS(r2,m_sqr_rpmax,_CMP_LT_OS);
if(AVX_TEST_COMPARISON(m_mask_left) == 0) {
  continue;
}

const AVX_FLOATS m_mask = AVX_BITWISE_AND(m_mask_left, AVX_COMPARE_FLOATS(r2, m_sqr_rpmin, _CMP_GE_OS));
if(AVX_TEST_COMPARISON(m_mask) == 0) {
  continue;
}
r2 = AVX_BLEND_FLOATS_WITH_MASK(m_sqr_rpmax, r2, m_mask);
m_mask_left = AVX_COMPARE_FLOATS(r2,m_sqr_rpmax,_CMP_LT_OS);
\end{lstlisting}

\subsubsection{Pair-counting for \texorpdfstring{\xirppi}{xi(rp,pi)}}\label{section:pair_counting_rp_pi}

\subsubsection{Pair-counting for \texorpdfstring{\wprp}{wp(rp)}}\label{section:pair_counting_wp}
Since the pair-counting in \wprp always assumes PERIODIC boundary conditions {\em and} only computes an auto-correlation, some more optimizations are possible in the
\wprp calculation. First, each individual cell in the \texttt{struct lattice} elements are sorted based on their \texttt{Z} arrays. The \texttt{X/Y} arrays are 
also re-ordered simultaneously. Once all the \texttt{Z} values inside a cell have been sorted, we can avoid double-counting the pairs by changing the \texttt{iiz loop}
to be:\\
\vspace{0.2in}
\texttt{for(int iiz={\bf 0};iiz<=zbin\_refine\_factor;iiz++)} 
\vspace{0.2in}
\\instead of\\
\vspace{0.2in}
\texttt{for(int iiz={\bf -zbin\_refine\_factor};iiz<=zbin\_refine\_factor;iiz++)} 
\vspace{0.2in}

Another advantage of sorting the \texttt{Z} values is earlier termination inside the actual calculation. The Listing~\ref{code:calculating_separations_wp} shows 
how to break early from the \texttt{j-loop}. Since the \texttt{z2} values are always stored in increasing order, if {\em all} values of \texttt{zdiff:=z2[j:j+NVEC-1]-z1}
are greater than \pimax, then none of the \texttt{zdiff} values in future iterations of the \texttt{j-loop} can be smaller than \pimax. When the code encounters 
such a scenario, it updates the \texttt{j} variable to \texttt{second->nelements} to ensure that the serial section of the code is not executed and breaks out the 
AVX \texttt{j-loop}. 
\begin{lstlisting}[label={code:calculating_separations_wp},caption={AVX intrinsics for calculating separations in \wprp and checking for early termination.}]
const AVX_FLOATS m_zdiff = AVX_SUBTRACT_FLOATS(m_z2,m_zpos);//z2[j:j+NVEC-1] - z1                                                                                                                                        
AVX_FLOATS m_mask_pimax = AVX_COMPARE_FLOATS(m_zdiff,m_pimax,_CMP_LT_OS);
const int test = AVX_TEST_COMPARISON(m_mask_pimax);
if(test == 0) {
  j = second->nelements;
  break;
}
\end{lstlisting}

>>>>

\usepackage{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{aas_macros}
\usepackage{rotating}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{color}
\usepackage{listings}
\usepackage{caption}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue,breaklinks=true]{hyperref}
\usepackage{dirtree}
\usepackage{adjustbox}
\usepackage{bytefield}
\usepackage{chngcntr}
\usepackage{alltt}

@@

%% Some of the short-cuts I use
\newcommand{\rmax}{\ensuremath{{r_{max}}}\xspace}
\newcommand{\xir}{\ensuremath{{\xi(r)}}\xspace}
\newcommand{\wprp}{\ensuremath{{w_p(r_p)}}\xspace}
\newcommand{\xirppi}{\ensuremath{{\xi(r_p,\pi)}}\xspace}
\newcommand{\pimax}{\ensuremath{{\pi_{\rm max}}}\xspace}
\newcommand{\hMpc}{\ensuremath{{h^{-1}Mpc}}\xspace}
\newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}
\newcommand{\clang}{{\texttt{clang}}\xspace}
\newcommand{\icc}{{\texttt{icc}}\xspace}
\newcommand{\gcc}{{\texttt{gcc}}\xspace}

@@

\title{Users Guide for Corrfunc}
%% Abuse the author macro to show all the info I want.
\author{Manodeep Sinha, \\
  Department of Physics \& Astronomy,\\
  Vanderbilt University,\\
  Nashville, TN 37235.\\
  \href{mailto:manodeep@gmail.com}{\texttt{manodeep@gmail.com}}
}

@@

\stdsection{Running the Codes}
The codes should run straight out of the box. Note, the \xir and \xirppi can compute both auto and cross-correlations with and without PERIODIC boundary conditions whereas the \wprp code 
only computes the auto-correlation with PERIODIC boundary conditions. Also, since both the \xir and \xirppi codes compute cross-correlations, pairs are {\em double-counted} in both 
\xir and \xirppi codes. You could, in theory, make the auto-correlation bit faster by only computing unique pairs (as is done in \wprp, see Section~\ref{section:pair_counting_wp} 
for details). However, since the total run-time will dominated by the cross-correlation (the number of randoms is typically an order of magnitude larger than the number 
of data points), I have not implemented those optimizations for the auto-correlation calculations. 
\subsection{Input File Formats}\label{section:formats}
The codes currently can handle these types of input data files:
\begin{itemize}
\item \texttt{ascii} -- White-space separated columns, format code is `a'.
\item \texttt{csv}   -- Comma-separated values, format code is `c'.
\item \texttt{fast-food} -- Fast-food, fortran binary format, format code is `f'.
The fast-food file format is described in detail in Section~\ref{section:fastfood}.
\end{itemize}
For the \texttt{ascii} and \texttt{csv} files, the code reads in the first three columns as the co-moving \texttt{X/Y/Z} arrays. Note, that more columns 
can be present but the code will ignore those columns.

@@

The \texttt{fast-food} file consists of a header:
\begin{lstlisting}[caption={The header format for \texttt{fast-food} files}]
int idat[5];
float fdat[9];
float znow;
\end{lstlisting}
For the purposes of these correlation function codes, the only useful quantity is \texttt{idat[1]} which contains \texttt{N} -- the number of particles 
in the data file.

@@

\vspace{0.2in}
\begin{minipage}[t]{\textwidth}
\centerline{\underline{\bfseries Byte-structure of the X/Y/Z arrays in a fast-food file}}
\vspace{0.2in}
\begin{minipage}[t]{0.5\textwidth}
\centering
\textbf{floats}\\[3ex]
\begin{bytefield}{4}
  \bitheader{0-3} \\
  \bitbox{4}{\texttt{N*4}} \\
  \wordbox{1}{X[0]} \\
  \wordbox{1}{X[1]} \\
  \wordbox[]{1}{$\vdots$} \\[1ex]
  \wordbox{1}{X[N-1]} \\
  \bitbox{4}{\texttt{N*4}}
\end{bytefield} 
\end{minipage}%
\begin{minipage}[t]{0.5\textwidth}
\centering
\textbf{doubles}\\[3ex]
\begin{bytefield}{8}
  \bitheader{0-7} \\
  \bitbox{4}{\texttt{N*8}} \\
  \wordbox{1}{X[0]} \\
  \wordbox{1}{X[1]} \\
  \wordbox[]{1}{$\vdots$} \\[1ex]
  \wordbox{1}{X[N-1]} \\
  \bitbox{4}{\texttt{N*8}}
\end{bytefield}
\end{minipage}
\end{minipage}
\subsection{Specifying the radial bins}\label{section:bins}
The codes were intentionaly designed to read in a set of (somewhat) arbitrary\footnote{I assume bins are non-overlapping} set of bin specifications from a file. This way, you can specify disjoint bin-edges as well as 
use $0.0$ as a bin edge (which would be impossible if $\log$ bins are assumed). The bins are to be specified in a (white-space separated) text file in this manner:
\begin{alltt}
r\_low[0]  r\_high[0]
r\_low[1]  r\_high[1]
\vdots
r\_low[nbins-1]  r\_high[nbins-1]
\end{alltt}
where, \texttt{r\_low[i]} and \texttt{r\_high[i]} are the left and right edges of the \texttt{i}'th bin respectively. The text files should contains as many lines as the number of 
bins desired. The \texttt{logbins} executable can be used to create such a text file containing $\log$ bins. The syntax for running \texttt{logbins} is:\\
\vspace{0.2in}
\texttt{./logbins rmin rmax nbins   > filename}
\vspace{0.2in}
\\An example of such a file with radial bins is the file \texttt{bins} in the \texttt{tests} directory. 
{\em Note, all of the three codes print the correlation function to \texttt{stdout} -- so be sure to redirect \texttt{stdout} to an output file. }

\subsection{Running \texorpdfstring{\xir}{xi(r)}}
To run the correlation function in 3-D, you will need to run the executable \texttt{DD} from either the \texttt{bin} or \texttt{xi\_of\_r} directory. Note, that 
\xir double-counts the pairs (as does \xirppi). The inputs to \texttt{DD} are:
\begin{itemize}
\item \texttt{file1}   -- the file name for the first file.
\item \texttt{format1} -- the file format for the first file. Options are \texttt{a,c,f} -- see Section~\ref{section:formats}.
\item \texttt{file2}   -- the file name for the second file.
\item \texttt{format2} -- the file format for the second file. Options are \texttt{a,c,f} -- see Section~\ref{section:formats}.
\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
\end{itemize}
Thus, the code can compute an auto-correlation (when \texttt{file1,format1} and \texttt{file2,format2} are identical) or a cross-correlation (when \texttt{file1} and \texttt{file2} are different). Out of the box, 
some sample \texttt{DD} call directory) can be:
\begin{itemize}
\item auto-correlation --  {\scriptsize \texttt{./DD  ../tests/data/gals\_Mr19.ff f ../tests/data/gals\_Mr19.ff f ../tests/bins 4 > Mr19\_output.DD}}
\item cross-correlation -- {\scriptsize \texttt{./DD ../tests/data/cmassmock\_Zspace.ff f ../tests/data/random\_Zspace.ff f ../tests/bins 4 > cmass\_output.DR}}
\end{itemize}
The output (printed to stdout) has \texttt{nbins} rows; each row contains the columns \texttt{<Npairs> <rpavg> <rmin> <rmax>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
Makefile option OUTPUT\_RPAVG has been enabled. Here, each line of the output represents the \texttt{i}'th radial bin.

\subsection{Running \texorpdfstring{\xirppi}{xi(rp,pi)}}
To run the correlation function in 2-D for \xirppi, you will need to run the executable \texttt{DDrppi} from either the \texttt{bin} or \texttt{xi\_rp\_pi} directory. Note, that 
\xirppi double-counts the pairs (as does \xir).The inputs 
to \texttt{DDrppi} are:
\begin{itemize}
\item \texttt{file1}   -- the file name for the first file.
\item \texttt{format1} -- the file format for the first file. Options are \texttt{a,c,f} -- see Section~\ref{section:formats}.
\item \texttt{file2}   -- the file name for the second file.
\item \texttt{format2} -- the file format for the second file. Options are \texttt{a,c,f} -- see Section~\ref{section:formats}.
\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
\item \texttt{pimax}   -- the maximum distance to consider in the $\pi$ direction. The code uses \texttt{Z} axis as the $\pi$ (line-of-sight) direction. 
\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
\end{itemize}
Thus, the code can compute an auto-correlation (when \texttt{file1,format1} and \texttt{file2,format2} are identical) or a cross-correlation (when \texttt{file1} and \texttt{file2} are different). Out of the box, 
some sample \texttt{DDrppi} call directory) can be:
\begin{itemize}
\item auto-correlation --  {\scriptsize \texttt{./DDrppi  ../tests/data/gals\_Mr19.ff f ../tests/data/gals\_Mr19.ff f ../tests/bins 40.0 4 > Mr19\_output\_rppi.DD}}
\item cross-correlation -- {\scriptsize \texttt{./DDrppi ../tests/data/cmassmock\_Zspace.ff f ../tests/data/random\_Zspace.ff f ../tests/bins 80.0 4 > cmass\_output\_rppi.DR}}
\end{itemize}
The output (printed to stdout) has \texttt{nbins} rows; each row contains the columns \texttt{<Npairs> <rpavg> <log(rmin)> <pi\_upper>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
Makefile option OUTPUT\_RPAVG has been enabled. The code bins in 1 Mpc/h in the $\pi$ direction by default. Therefore, 
the total number of bins in the output file will be \texttt{nbins$\times$\pimax}, where each radial bin is further broken into \pimax bins along the 
$\pi$ direction.

@@

\subsection{Running \texorpdfstring{\wprp}{wp(rp)}}
To run the projected correlation function, you will need to run the executable \texttt{wp} from either the \texttt{bin} or \texttt{wp} directory. The inputs 
to \texttt{wp} are:
\begin{itemize}
\item \texttt{boxsize} -- the boxsize for the periodic box. 
\item \texttt{file}    -- file name.
\item \texttt{format}  -- file format. Options are \texttt{a,c,f} -- see Section~\ref{section:formats}.
\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
\item \texttt{pimax}   -- the maximum distance to consider in the $\pi$ direction. The code uses \texttt{Z} axis as the $\pi$ (line-of-sight) direction. 
\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
\end{itemize}
The \texttt{wp} code only computes an auto-correlation with PERIODIC boundary conditions, irrespective of the options set in \texttt{common.mk}. 
\begin{itemize}
\item {\scriptsize \texttt{./wp  420.0 ../tests/data/gals\_Mr19.ff f ../tests/bins 40.0 4 > Mr19\_output.wp}}
\end{itemize}
The output (printed to stdout) has \texttt{nbins} rows; each row contains the columns \texttt{<wp> <rpavg> <rmin> <rmax> <Npairs>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
Makefile option OUTPUT\_RPAVG has been enabled.

@@

\subsection{Partitioning the Particles based on \texorpdfstring{\rmax}{rmax}}
We need to compute pairwise distances to get the correlation function. A naive implementation of a correlation function would compute {\em all possible} 
pairwise separations with a complexity $\mathcal{O}(N^2)$. However, for almost all correlation functions, we are only interested in separations less than 
a certain \rmax, where \rmax is much smaller than the domain of the point distribution itself. We can then immediately see a way to prune pairs that can not
{\em possibly} be within \rmax. If we impose a 3-d grid, with cell-size \rmax, then two points separated by more than one cell size (\rmax) in any one 
dimension can not be within \rmax of each other (see Fig.~\ref{fig:grid} for a 2-D representation). Thus, given one point which is the target galaxy and a grid with cell-size \rmax, 
immediately allows us to prune {\em all} of the points that are not within 1 cell offset in each dimension. However, even with this pruning, the actual 
implementation of the algorithm matters. For instance, the non-AVX version of the codes tend to run $2-3\times$ slower than the AVX version. 
\begin{figure}
\centering
\includegraphics[width=\textwidth,clip=true]{tikz_grid}
\caption{A 2-D grid showing the bin-lattice partitioning scheme. The bigger square show the entire
domain, the red circles show a random distribution of 100 particles. Let's say we want to compute all pairs
for the target blue point, then we would only have to consider red points that are within one cell (the dark shaded region).
A circle with radius \rmax is also drawn to shown the actual pairs that will eventually count in the correlation function.}
\label{fig:grid}
\end{figure}

@@

By wrapping the elements if the first data-set, we can avoid the wrapping operations in the \texttt{j-loop} over all particles in \texttt{second}. 
\begin{lstlisting}[label={code:looping_over_second_particles},caption={AVX intrinsics for looping over all particles in the \texttt{second} cell. PERIODIC boundary conditions have 
already been accounted for in \texttt{x1pos,y1pos,z1pos} variables.}]
const AVX_FLOATS m_x1pos = AVX_SET_FLOAT(x1pos);
const AVX_FLOATS m_y1pos = AVX_SET_FLOAT(y1pos);
const AVX_FLOATS m_z1pos = AVX_SET_FLOAT(z1pos);

@@

The three codes \xir, \xirppi and \wprp diverge somewhat after this point. For \xir, we need to calculate the full 3-D separation, whereas for \xirppi and \wprp 
the separation is the projected distance. I will discuss further implementations in the following sub-sections dedicated to each code. 
\subsubsection{Pair-counting for \texorpdfstring{\xir}{xi(r)}}\label{section:pair_counting_r}
Pair-counting for \xir is the most straight-forward. The squared 3-D separation, \texttt{r2}, is simply the sum of the squared differences in 
each \texttt{X/Y/Z} dimensions. 
\begin{lstlisting}[label={code:r2_xi},caption={Calculating squared separations in \xir.}]
const AVX_FLOATS m_xdiff_sqr = AVX_SQUARE_FLOAT(m_xdiff);
const AVX_FLOATS m_ydiff_sqr = AVX_SQUARE_FLOAT(m_ydiff);
const AVX_FLOATS m_zdiff_sqr = AVX_SQUARE_FLOAT(m_zdiff);
const AVX_FLOATS m_xydiff_sqr_sum = AVX_ADD_FLOATS(m_xdiff_sqr,m_ydiff_sqr);
AVX_FLOATS r2 = AVX_ADD_FLOATS(m_zdiff_sqr,m_xydiff_sqr_sum);
\end{lstlisting}
Once all the \texttt{NVEC} separations have been computed, we use bit-masks to check if any separations fall 
within the range \texttt{sqr\_rpmin} and \texttt{sqr\_rpmax}. If not, we \texttt{continue} with the \texttt{j-loop}. 
\begin{lstlisting}[label={code:masks_xi},caption={Bit-masks in \xir.}]
m_mask_left = AVX_COMPARE_FLOATS(r2,m_sqr_rpmax,_CMP_LT_OS);
if(AVX_TEST_COMPARISON(m_mask_left) == 0) {
  continue;
}

const AVX_FLOATS m_mask = AVX_BITWISE_AND(m_mask_left, AVX_COMPARE_FLOATS(r2, m_sqr_rpmin, _CMP_GE_OS));
if(AVX_TEST_COMPARISON(m_mask) == 0) {
  continue;
}
r2 = AVX_BLEND_FLOATS_WITH_MASK(m_sqr_rpmax, r2, m_mask);
m_mask_left = AVX_COMPARE_FLOATS(r2,m_sqr_rpmax,_CMP_LT_OS);
\end{lstlisting}
If the code reaches past this bit-mask section, then at least one separation is within range. In that case, we use the 
code in Listing~\ref{code:updating_histogram} to update the \texttt{npairs} array (and the \texttt{rpavg} array if 
OUTPUT\_RPAVG is enabled).

\subsubsection{Pair-separations for the projected correlation functions \texorpdfstring{\xirppi,\wprp}{xi(rp,pi), wp(rp)}}
Listing~\ref{code:r2_projected} shows the squared distance calculation for \xirppi and \wprp. \texttt{r2} is simply 
the projected separation in the \texttt{X-Y} plane.
\begin{lstlisting}[label={code:r2_projected},caption={Calculating squared separations in \xirppi and \wprp.}]
const AVX_FLOATS m_xdiff_sqr = AVX_SQUARE_FLOAT(m_xdiff);
const AVX_FLOATS m_ydiff_sqr = AVX_SQUARE_FLOAT(m_ydiff);
AVX_FLOATS r2 = AVX_ADD_FLOATS(m_xdiff_sqr,m_ydiff_sqr);
\end{lstlisting}

\subsubsection{Pair-counting for \texorpdfstring{\xirppi}{xi(rp,pi)}}\label{section:pair_counting_rp_pi}
\begin{lstlisting}[label={code:masks_xirppi},caption={Bit-masks in \xirppi.}]
const AVX_FLOATS m_mask_pimax = AVX_COMPARE_FLOATS(m_zdiff,m_pimax,_CMP_LT_OS);
const int test = AVX_TEST_COMPARISON(m_mask_pimax);
if(test == 0) {
  continue;
}

const AVX_FLOATS m1 = AVX_COMPARE_FLOATS(r2,m_sqr_rpmin,_CMP_GE_OS);
r2 = AVX_BLEND_FLOATS_WITH_MASK(m_sqr_rpmax,r2,m_mask_pimax);

m_mask_left = AVX_COMPARE_FLOATS(r2,m_sqr_rpmax,_CMP_LT_OS);
const AVX_FLOATS m_mask = AVX_BITWISE_AND(m1,m_mask_left);
int test1 = AVX_TEST_COMPARISON(m_mask);
if(test1 == 0) {
  continue;
}
m_zdiff = AVX_BLEND_FLOATS_WITH_MASK(m_pimax, m_zdiff, m_mask);
#ifdef OUTPUT_RPAVG
union_mDperp.m_Dperp = AVX_SQRT_FLOAT(r2);
#endif
union_pibin.m_ibin = AVX_TRUNCATE_FLOAT_TO_INT(AVX_MULTIPLY_FLOATS(m_zdiff,m_inv_dpi));
\end{lstlisting}
In the \xirppi code, we have to update a two dimensional \texttt{npairs($r_p$,$\pi$)} array. This means we can not directly update 
the \texttt{npairs} matrix and instead have to use a separate loop (this loop is typically only invoked for \xir and \wprp 
when OUTPUT\_RPAVG is enabled). As a result of this extra loop, \xirppi is $2-3\times$ slower than \xir and \wprp.

\begin{lstlisting}[label={code:updating_histogram_xirppi},caption={Updating the \texttt{npairs} matrix in \xirppi.}]
for(int jj=0;jj<NVEC;jj++) {
  int rpbin = union_rpbin.ibin[jj];
  int pibin = union_pibin.ibin[jj];
  int ibin = rpbin*(npibin+1) + pibin;
  npairs[ibin]++;
#ifdef OUTPUT_RPAVG
  rpavg [ibin] += union_mDperp.Dperp[jj];
#endif
}
\end{lstlisting}
